Argennon: A Scalable Smart Contract Platform

### 2021-06-24 tex/sections/consensus.tex
Initially fix the consensus problem
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

Argennon has two primary certificate committees: the committee of \emph{delegates} and the committee
of \emph{validators}. A block is considered final only after it gets a certificate from both of these committees.

The committee of delegates generates a certificate for every block of the Argennon blockchain, and the
committee of validators generates a certificate every $m$ blocks. The validators' committee will
certify a block only if it has already been certified by the committee of delegates.

@@

The committee of delegates is a small committee of trusted delegates, elected by Argennon users either through the
Argennon Decentralized Autonomous Governance system (ADAGs\footnote{pronounced \textipa{/eI-dagz/}.}) or by emergency
voting during the recovery protocol. At the start of the Argennon mainnet, this committee will have
five members, and later its size could be changed by the ADAGs in a procedure described
in Section~\ref{sec:governance}.

@@

The Argennon protocol calculates a stake value for every account, which is an estimate of a user's stake in the
system. This value is measured in ARGs. The committee of validators is composed of
\textbf{all} accounts whose stake value is higher than \texttt{minValidatorsStake} threshold. This
threshold is determined by the ADAGs, but it can never be higher
than $100$ ARGs.

Every member of the validators committee has a status which can be either \texttt{online} or \texttt{offline}.
This status is stored in the ARG smart contract, and it can be changed through a method invocation
from this smart contract. When an account sets its status to \texttt{offline}, it receives a small reward, and
it can not change it back to \texttt{online} for \texttt{statusCoolDown} number of blocks.

@@

A block certificate issued by some members of the validators committee is considered valid, if according to
the staking database of the previous \textbf{certified block by the validators' committee}, we have:
\begin{itemize}
    \item The total stake of \texttt{online} accounts is higher than \texttt{minOnlineStake} percent of the
    total stake of all accounts. This threshold can be changed by the ADAGs, but it can never be lower
    than 70 percent.
    \item All signers of the certificate have \texttt{online} status.
    \item The sum of stake values of the certificate signers is higher than 80 percent of the total stake
    of accounts that have \texttt{online} status.
\end{itemize}

If the total online stake goes lower than \texttt{minOnlineStake} threshold, nodes initiate the recovery
consensus protocol.

As we mentioned before, every $m$ blocks of the Argennon blockchain needs a certificate from the committee
of validators. This means that assuming the last certified block by the validators is valid, for issuing the block
certificate, the committee needs to conditionally validate $m$ blocks. We call these $m$ blocks the
validation set. To reduce resource usage of validators' nodes, the protocol requires honest nodes to only
conditionally validate a subset of the validation set. Validators that have a stake lower than
average should only validate one block in expectation.

More precisely, to decide about signing the block certificate, a honest member of the validators'
committee iterates over every block of the validation set, and selects a block for validation with
probability $p_u$. In this way, a member of the validators' committee will select $p_{u}m$ blocks in expectation.
Then, the honest member conditionally validates the selected blocks. If the member has selected more than one block,
he may validate the selected blocks in parallel.
If all checked blocks are valid, the validator issues an "accept"
signature, otherwise he issues a "reject" signature.

One recommended formula for calculating the probability of selecting a block by user $u$ for validation is:
\begin{equation}
    p_u=\max\left(\frac{s_u}{m\mu},\ \frac{1}{m}\right),\label{eq:block-selection}
\end{equation}
where $\mu$ is the average stake of committee members and $s_u$ is the stake of user $u$.

The value of $m$ is determined by the ADAGs. but it can never be higher than $100$. This way, it is guaranteed
that on average, any block of the Argennon blockchain is validated by at least $0.005$ of the total ARG supply.

If the committee of validators, at any point issues a "reject" certificate for a block, the network will initiate the
recovery protocol.

>>>>

There is two primary types of certificate committees in Argennon: the committee of \emph{delegates} and the committee
of \emph{validators}. Argennon has one committee of delegates and $m$ committees of validators.

The committee of delegates generates a certificate for every block of the Argennon blockchain, and each
committee of validators generates a certificate every $m$ blocks. A validators' committee will
certify a block only if it has already been certified by the committee of delegates and every committee
generates a certificate for a different block. A block is considered
final only after it gets a certificate from both of these committees.

@@

The committee of delegates is a small committee of trusted delegates, elected by Argennon users either through the
Argennon Decentralized Autonomous Governance system (ADAGs\footnote{pronounced \textipa{/eI-dagz/}.}) or by emergency
voting during the recovery protocol. At the start of the Argennon mainnet, this committee will have
five members, and later its size could be changed by the ADAGs in a procedure described
in Section~\ref{ch:governance}.

@@

The Argennon protocol calculates a stake value for every account, which is an estimate of a user's stake in the
system. This value is measured in ARGs. Every $n$ block randomly $m$ committees are selected from
accounts whose stake value is higher than \texttt{minValidatorsStake} threshold. This
threshold is determined by the ADAGs, but it can never be higher than $100$ ARGs.
These committees are chosen in a
way that the total stake of members of each committee is approximately equal and every account is a member of at
least one committee.

Every member of a validators committee has a status which can be either \texttt{online} or \texttt{offline}.
This status is stored in the ARG smart contract, and it can be changed through a method invocation
from this smart contract. When an account sets its status to \texttt{offline}, it receives a small reward, and
it can not change it back to \texttt{online} for \texttt{statusCoolDown} number of blocks.

@@

A block certificate issued by some members of a validators committee is considered valid, if according to
the staking database of the previous \textbf{block certified by the same committee}, we have:
\begin{itemize}
    \item The total stake of \texttt{online} members is higher than \texttt{minOnlineStake} percent of the
    total stake of all members. This threshold can be changed by the ADAGs, but it can never be lower
    than 70 percent.
    \item All signers of the certificate have \texttt{online} status.
    \item The sum of stake values of the certificate signers is higher than 80 percent of the total stake
    of committee members that have \texttt{online} status.
\end{itemize}

If the total online stake of a committee goes lower than \texttt{minOnlineStake} threshold, nodes initiate the recovery
consensus protocol.

As we mentioned before, every $m$ blocks of the Argennon blockchain needs a certificate from a specific committee
of validators. To decide about signing the block certificate, an honest member of the validators'
committee checks the conditional validity of the committee block, and if the block is valid he issues an "accept"
signature, otherwise he issues a "reject" signature.

The value of $m$ is determined by the ADAGs. but it can never be higher than $25$. This way, it is guaranteed
that on average, any block of the Argennon blockchain is validated by at least $0.02$ of the total ARG supply.

If a committee of validators, at any point issues a "reject" certificate for a block, the network will initiate the
recovery protocol.
